---
title: "Lab4"
author: "Molly Carrick"
date: "10/7/2020"
output: html_document
---

```{r}
library(readxl)
library(tidyverse)
defense <- read_excel('cyclonesFootball2019.xlsx', sheet='Defensive')
offense <- read_excel('cyclonesFootball2019.xlsx', sheet='Offensive')
biography <- read_excel('cyclonesFootball2019.xlsx', sheet='Biography')
```



```{r}
defClean<-defense%>%
  rename(Opponent=Opponent_Opponent)%>%
  mutate(Opponent=factor(Opponent))%>%
  mutate(Name=factor(Name))%>%
  mutate(Tackles_Solo=as.numeric(Tackles_Solo))%>%
  mutate(Tackles_ASST=as.numeric(Tackles_ASST))%>%
  mutate(Tackles_TFL=as.numeric(Tackles_TFL))%>%
  mutate(Tackles_Sack=as.numeric(Tackles_Sack))%>%
  mutate(Turnover_FF=as.numeric(Turnover_FF))%>%
  mutate(Turnover_FR=as.numeric(Turnover_FR))%>%
  mutate(Turnover_INT=as.numeric(Turnover_INT))%>%
  mutate(Pass_QBH=as.numeric(Pass_QBH))%>%
  mutate(Pass_PB=as.numeric(Pass_PB))

offClean<-offense%>%
  rename(Opponent=Opponent_Opponent)%>%
  mutate(Opponent=factor(Opponent))%>%
  mutate(Name=factor(Name))%>%
  mutate(Rushing_ATT=as.numeric(Rushing_ATT))%>%
  mutate(Rushing_YDS=as.numeric(Rushing_YDS))%>%
  mutate(Rushing_TD=as.numeric(Rushing_TD))%>%
  mutate(Receiving_REC=as.numeric(Receiving_REC))%>%
  mutate(Receiving_YDS=as.numeric(Receiving_YDS))%>%
  mutate(Receiving_TD=as.numeric(Receiving_TD))%>%
  mutate(`Passing_CMP-ATT`=as.numeric(`Passing_CMP-ATT`))%>%
  mutate(Passing_YDS=as.numeric(Passing_YDS))%>%
  mutate(Passing_TD=as.numeric(Passing_TD))%>%
  mutate(Passing_INT=as.numeric(Passing_INT))


bioClean<-biography%>%
  mutate(Name=factor(Name))%>%
  separate(Height, into=c('feet','inches'), sep='-')%>%
  mutate(feet=as.numeric(feet))%>%
  mutate(inches=as.numeric(inches))

bioClean$feet<-bioClean$feet * 12
bioClean$Height<-bioClean$feet+bioClean$inches
bioClean<-select(bioClean,-feet,-inches)

str(defClean)
str(offClean)
str(bioClean) 
```



```{r}
defClean2<-defClean%>%
  pivot_longer(Tackles_Solo:Pass_PB, names_to="type", values_to='stat')

str(defClean2) ## Compare Distributions/scatter plot
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
bioClean<-bioClean%>%
  separate(Hometown, into=c('City','State'), sep=',')


##How many players are there from each state?
```



Part 3
Find which states do the ISU defensive and offensive players come from. Create an appropriate summary.

```{r}
playerState<-bioClean%>%
  select(Name,State)

originDef<-playerState%>%
  semi_join(defClean, by="Name")

ggplot(originDef, aes(x=State))+geom_bar()+ggtitle("Defensive Player Origin")

originOff<-playerState%>%
  semi_join(offClean, by="Name")

ggplot(originOff, aes(x=State))+geom_bar()+ggtitle("Offensive Player Origin")

str(originDef)
```
```{r}

str(originOff)
```

How does the performance of quarterback Brock Purdy relate to the performance of the ISU team?

```{r}
count(originDef, 'State')
count(originOff, 'State')
```
```{r}

try<-playerState%>%
  left_join(defClean, by="Name")

try<-try%>%
  inner_join(playerState, by="Name")
```


```{r}

playerOff<-offClean%>%
  select(Name)

class(playerOff)
levels(playerOff$Name)

df<-data.frame(Name=levels(playerOff$Name))


dataframe<-df%>%
  left_join(playerState, by="Name")

ggplot(dataframe, aes(x=State))+geom_bar()+ggtitle("Offensive Player Origin")
```


How does the performance of quarterback Brock Purdy relate to the performance of the ISU team?

```{r}


```





